---
title: "PMM Model"
author: "Chandan Basavaraju"
date: "Monday, November 17, 2014"
output: html_document
---

The files for PMM-training.csv and PMM-testing.csv were downloaded and then the following operations were performed:

Data Cleaning: 

Step 1: Load in Excel and perform cleaning up of #DIV/0! occurences in the pml-training.csv and pml-testing.csv. The output from these are pml-training_data.csv and pml-testing_data.csv respectively


Step 2: Load the data in R 

```{r}
setwd("c:\\data")
training<-read.csv("pml-training_data.csv")
testing<-read.csv("pml-testing_data.csv")
```

Step 3: Removed all columns which contained a majority of NA ( > 50% )
```{r}
     trainingpruned<-training[,colSums(is.na(training))< nrow(training)/2]
     testingpruned<-testing[,colSums(is.na(testing))< nrow(testing)/2]

```
Step 4:  Replaced remaining occurences of NA with 0
```{r}
     trainingpruned[is.na(trainingpruned)]<-0
     testingpruned[is.na(testingpruned)]<-0
```

Step 5: Removed the Redundant date columns (3,5) and the first column which was for SI
```{r}
trainingpruned<-trainingpruned[,c(-1,-3,-5)]
testingpruned<-testingpruned[,c(-1,-3,-5)]
```

Step 6: Split the data into training data into subsets of trainingset and validationset:
```{r}
library(caret)
intrain<-createDataPartition(trainingpruned$classe, p=0.8, list=FALSE)
trainingset<-trainingpruned[intrain,]
validationset<-trainingpruned[-intrain,]
```

Step 7: Build a model

The model was build using the following code: 
```{r}
gbmmodel<-train(trainingset$classe~.,data=trainingset, method="gbm")
```

Step 8: Cross Validate using the validationSet
```{r}
gbmvalidationpredict<-predict(gbmmodel, validationset, type="raw")
table(validationset$classe,gbmvalidationpredict)
```

Step 9: Calculate out of sample error
```{r}
t<-table(validationset$classe,gbmvalidationpredict)
denominator<-sum(t)
diag(t)<-0
numerator <-sum(t)
outofsampleerror = numerator*100/denominator
outofsampleerror
```

Step 8: Predict the testing set

```{r}
gbmTestingpredict<-predict(gbmmodel, testingpruned, type="raw")
gbmTestingpredict
```
